---
title: "PML course project"
author: "Tsitko"
date: "Monday, May 18, 2015"
output: html_document
---


---
Practical Machine Learning Course Project by Denis Tsitko (18.05.2015)
---

---
Librarys
---



```{r}
library(rpart)
library(rpart.plot)
library(randomForest)
library(caret)
library(corrplot)
```


---
Downloading and reading data
---


```{r}
if (!file.exists("train.csv")) {
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile="train.csv")
}
if (!file.exists("test.csv")) {
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile="test.csv")
}

train<-read.csv("train.csv")
test<-read.csv("test.csv")
```

---
Cleaning data
---

Lets look at the data dimentions:

```{r echo=TRUE}
dim(train)
dim(test)
```

So there are 160 variables in each data set. And there is 19622 observations for training set and 20 oservations for test set.

Lets leave only variables with all complete cases:

```{r}
train <- train[, colSums(is.na(train)) == 0] 
test <- test[, colSums(is.na(test)) == 0] 
class <- train$classe
train2Rm <- grepl("^X|timestamp|window", names(train))
train <- train[, !train2Rm]
trainFin <- train[, sapply(train, is.numeric)]
trainFin$classe <- class
test2Rm <- grepl("^X|timestamp|window", names(test))
test <- test[, !test2Rm]
testFin <- test[, sapply(test, is.numeric)]
```

Lets look at the cleaned data dimentions:

```{r echo=TRUE}
dim(trainFin)
dim(testFin)
```

So the nomber of observations are the same, but the number of variables for each set is now 53.

---
Creating preTrain and preTest sets
---

```{r}
set.seed(123) 
inTrain <- createDataPartition(trainFin$classe, p=0.60, list=F)
preTrain <- trainFin[inTrain, ]
preTest <- trainFin[-inTrain, ]
```

---
Creating the model
---

Here we will use a simple randomForest model training (using train from caret with defaults) for our data.

```{r echo=TRUE}
set.seed(123)
trC <- trainControl(method="cv", 3)
md<-train(classe~.,data=preTrain, trControl=trC,ntree=200)
md
```

Now we predicting for preTets set:

```{r echo=TRUE}
res <- predict(md, preTest)
confusionMatrix(preTest$classe, res)
```


```{r echo=TRUE}
accuracy <- postResample(res, preTest$classe)
accuracy
```

So we have accuracy of about 99%.

Lets now update our model with preTest data and make a prediction for our testFin data:

```{r echo=TRUE}
md<-update(md,data=preTest)
prediction <- predict(md, testFin[, -length(testFin[1,])])
prediction
```

---
Plots
---

Lets plot correlation matrix:

```{r echo=TRUE}
corrMat <- cor(trainFin[, -length(trainFin[1,])])
corrplot(corrMat, method="color")
```

Lets plot decision tree:

```{r echo=TRUE}
tree <- rpart(classe ~ ., data=trainFin, method="class")
prp(tree) 
```